### 2017-05-31 letter-editor.tex
Added letter to editor tex file

\documentclass{letter}
\begin{document}

To \\
Editorial board, AGU Space Weather,\\

\hfill

\end{document}